Extracting sunbeams from cucumbers: Reproducibility as a necessity for the future of archaeological science

### 2025-01-28 paper.qmd
typing

The most striking change over time is in the increase in the number of authors, which has the highest r-squared value of these metrics. One interesting detail evident in @fig-change-over-time is the increase in the range of diversity of references after about 2010. This may be due to some broader changes in academic publishing around this time, such as moves to digital-first continuous publishing, new journals appearing (e.g. _Archaeological and Anthropological Sciences_ in 2009 and _Journal of Island & Coastal Archaeology_ in 2010), and non-archaeology journals becoming more relevant to archaeologists. For example, *PLOS ONE* received its first impact factor in 2010 and in 2011 Nature's *Scientific Reports* began publishing [@malashichev2017open].

@@

The PCA results in panel G show of @fig-variation-by-journal that PC1 captures most of the variance in the metrics (`r variance_explained[1]`%) and is reasonable proxy for the hard-soft spectrum, with *Journal of Cultural Heritage* representing the hard extreme on the right and *Journal of Archaeological Research* representing the soft extreme on the left. The distribution of PC1 values is skewed left, with most of the journals concentrated at the harder end of the spectrum. The distribution of PC2 values reveals additional structure to the data and can be roughly separated into generalist journals in the negative range (e.g. _American Antiquity_, _Antiquity_, _Advances in Archaeological Practice_), and more specialised journals in the positive range (e.g. _Environmental Archaeology_, _Geoarchaeology_, _Archaeologial Research in Asia_, _Journal of Island and Coastal Archaeology_). The _Journal of Archaeological Science_ sits about midway between these two groups, reflecting its relevance to specialised communities of practice in archaeology, as well as general

## Dicsussion

@@

>>>>

The most striking change over time is in the increase in the number of authors, which has the highest r-squared value of these metrics. One interesting detail evident in @fig-change-over-time is the increase in the range of diversity of references after about 2010. This may be due to some broader changes in academic publishing around this time, such as moves to digital-first continuous publishing, new journals appearing (e.g. *Archaeological and Anthropological Sciences* in 2009 and *Journal of Island & Coastal Archaeology* in 2010), and non-archaeology journals becoming more relevant to archaeologists. For example, *PLOS ONE* received its first impact factor in 2010 and in 2011 Nature's *Scientific Reports* began publishing [@malashichev2017open].

@@

The PCA results in panel G show of @fig-variation-by-journal that PC1 captures most of the variance in the metrics (`r variance_explained[1]`%) and is reasonable proxy for the hard-soft spectrum, with *Journal of Cultural Heritage* representing the hard extreme on the right and *Journal of Archaeological Research* representing the soft extreme on the left. The distribution of PC1 values is skewed left, with most of the journals concentrated at the harder end of the spectrum. The distribution of PC2 values reveals additional structure to the data and can be roughly separated into generalist journals in the negative range (e.g. *American Antiquity*, *Antiquity*, *Advances in Archaeological Practice*), and more specialised journals in the positive range (e.g. *Environmental Archaeology*, *Geoarchaeology*, *Archaeologial Research in Asia*, *Journal of Island and Coastal Archaeology*). The *Journal of Archaeological Science* sits about midway between these two groups, reflecting its relevance to specialised communities of practice in archaeology, as well as general

## Discussion

@@

With many disciplines making increasing use of computationally intensive analyses in recent years there has been renewed interest in reproducibility [@levequeReproducibleResearchScientific2012]. In part, this is because computationally intensive research is difficult to communicate within the constraints of the methods section of a traditional journal article — the reader needs also needs the computer code written by the original authors, not just the article text. There is also the broader context of rising pressure to publish in prestigious journals and intense competition for funds that create strong incentives for malpractice in research [@edwards2017academic]. These two factors have led to widespread concerns of a reproducibility crisis in many fields [@Baker2016]. Estimates of the extent of this problem come from both empirical and computational analyses. Empirical replications of 100 studies published in three psychology journals found that 36% of replications had statistically significant results, compared to 96% of the original studies [@OpenScienceCollaboration2015]. Similar empirical replications of social science studies and experiemental economics studies successfully replicated 61% and 62% of their target studies respectively [@camererEvaluatingReplicabilitySocial2018; @camererEvaluatingReplicabilityLaboratory2016]. An attempt at reproducing the computational results of 204 papers in *Science* resulted in success for reproduce the findings for 26% [@stoddenEmpiricalAnalysisJournal2018].
